## Analysing DNA damage/repair in response to different doses and types of radiation.
## In human samples. 
## Eloise Pariset, USRA/NASA Ames, January 2020
## BNL18B: Fe OK, excluded bad plates / Ar and Si only for fold difference, need to be re-imaged for absolute values
## BNL19A: Only Sets A here, all OK

## To be expanded (in the same file) by adding revised data on BNL18B + BNL19A sets B + all BNL19C
## To be expanded (in a separate file) by adding data on cell death, CellROX intensity and cell number


```{r}
require('lattice')
require('ggplot2')
require('polynom')
require('corrplot')
require('ggpubr')
require('plyr')
require('reshape2')
require('cowplot')
require('stringr')
```

```{r}
setwd("/Users/epariset/Desktop/R_Code-work") ## Change as necessary: this is the main directory
theme_set(theme_bw(base_size = 12)) ## For prettier graphs
theme_update(plot.title = element_text(hjust = 0.5))
```

```{r}
## BNL18B:
ar_plate_nums_18B = c(seq(379,402))
ar_plate_count_18B = length(ar_plate_nums_18B)
si_plate_nums_18B = c(seq(331,354))
si_plate_count_18B = length(si_plate_nums_18B)
fe_plate_nums_18B = c(seq(355,362),365,366,seq(368,371),374,seq(376,378))
fe_plate_count_18B = length(fe_plate_nums_18B)
si_fe_ar_plate_nums_18B = c(si_plate_nums_18B, fe_plate_nums_18B, ar_plate_nums_18B)
si_fe_ar_plate_count_18B = length(si_fe_ar_plate_nums_18B)

## BNL19A:
gamma_plate_nums_19A = c(433, 435, 437, 439, 441, 444, 445, 447, 449, 451, 453, 455, 457, 459, 461, 463, 465, 467, 469, 471, 473, 475, 477)
gamma_plate_count_19A = length(gamma_plate_nums_19A)
ar_plate_nums_19A = c(481, 483, 485, 487, 489, 491, 493, 495, 497, 499, 503, 505, 507, 509, 511, 513, 515, 517, 519, 521, 523, 525)
ar_plate_count_19A = length(ar_plate_nums_19A)
si_plate_nums_19A = c(577, 579, 581, 583, 585, 587, 589, 591, 593, 595, 597, 599, 601, 603, 607, 609, 611, 615, 617, 619, 621, 623)
si_plate_count_19A = length(si_plate_nums_19A)
fe_plate_nums_19A = c(529, 531, 533, 535, 537, 539, 541, 543, 545, 547, 549, 551, 553, 555, 557, 559, 561, 563, 565, 567, 569, 571, 573)
fe_plate_count_19A = length(fe_plate_nums_19A)
gamma_si_fe_ar_plate_nums_19A = c(gamma_plate_nums_19A, si_plate_nums_19A, fe_plate_nums_19A, ar_plate_nums_19A)
gamma_si_fe_ar_plate_count_19A = length(gamma_si_fe_ar_plate_nums_19A)
```

```{r}
## Read all metadata
subject_meta = read.csv('/Users/epariset/Desktop/R_Code-work/Sample_metadata.csv', sep = ",", header = TRUE)
colnames(subject_meta)[colnames(subject_meta)=="ï..Sample_ID"]<-"Sample_ID"
subject_ids = unique(subject_meta$Sample_ID)
num_subj = length(subject_ids)
plate_meta = read.csv('/Users/epariset/Desktop/R_Code-work/Plate_metadata.csv', sep = ",", header = TRUE)
plate_ids = unique(plate_meta$Plate_number)
num_plate = length(plate_ids)
```

```{r}
## Read first file

## BNL18B:
foci_si_first_18B = data.frame(Plate = si_plate_nums_18B[1], Plate_well = paste(si_plate_nums_18B[1], read.table(paste("/Users/epariset/Desktop/Processed-all/BNL18B-ProcessedFile/Si-BNL_18B/average_well_P", si_plate_nums_18B[1],'.txt', sep=""), sep="\t", header = TRUE)$Well, sep = "_"), read.table(paste("/Users/epariset/Desktop/Processed-all/BNL18B-ProcessedFile/Si-BNL_18B/average_well_P", si_plate_nums_18B[1],'.txt', sep=""), sep="\t", header = TRUE))

si_table_18B=foci_si_first_18B

foci_ar_first_18B = data.frame(Plate = ar_plate_nums_18B[1], Plate_well = paste(ar_plate_nums_18B[1], read.table(paste("/Users/epariset/Desktop/Processed-all/BNL18B-ProcessedFile/Ar-BNL_18B/average_well_P", ar_plate_nums_18B[1],'.txt', sep=""), sep="\t", header = TRUE)$Well, sep = "_"), read.table(paste("/Users/epariset/Desktop/Processed-all/BNL18B-ProcessedFile/Ar-BNL_18B/average_well_P", ar_plate_nums_18B[1],'.txt', sep=""), sep="\t", header = TRUE))

ar_table_18B=foci_ar_first_18B

foci_fe_first_18B = data.frame(Plate = fe_plate_nums_18B[1], Plate_well = paste(fe_plate_nums_18B[1], read.table(paste("/Users/epariset/Desktop/Processed-all/BNL18B-ProcessedFile/Fe-BNL_18B/average_well_P", fe_plate_nums_18B[1],'.txt', sep=""), sep="\t", header = TRUE)$Well, sep = "_"), read.table(paste("/Users/epariset/Desktop/Processed-all/BNL18B-ProcessedFile/Fe-BNL_18B/average_well_P", fe_plate_nums_18B[1],'.txt', sep=""), sep="\t", header = TRUE))

fe_table_18B=foci_fe_first_18B

## BNL19A:
foci_gamma_first_19A = data.frame(Plate = gamma_plate_nums_19A[1], Plate_well = paste(gamma_plate_nums_19A[1], read.table(paste("/Users/epariset/Desktop/Processed-all/BNL19A-ProcessedFiles-Threshold3000/average_well_P", gamma_plate_nums_19A[1],'.txt', sep=""), sep="\t", header = TRUE)$Well, sep = "_"), read.table(paste("/Users/epariset/Desktop/Processed-all/BNL19A-ProcessedFiles-Threshold3000/average_well_P", gamma_plate_nums_19A[1],'.txt', sep=""), sep="\t", header = TRUE))

gamma_table_19A=foci_gamma_first_19A

foci_si_first_19A = data.frame(Plate = si_plate_nums_19A[1], Plate_well = paste(si_plate_nums_19A[1], read.table(paste("/Users/epariset/Desktop/Processed-all/BNL19A-ProcessedFiles-Threshold3000/average_well_P", si_plate_nums_19A[1],'.txt', sep=""), sep="\t", header = TRUE)$Well, sep = "_"), read.table(paste("/Users/epariset/Desktop/Processed-all/BNL19A-ProcessedFiles-Threshold3000/average_well_P", si_plate_nums_19A[1],'.txt', sep=""), sep="\t", header = TRUE))

si_table_19A=foci_si_first_19A

foci_ar_first_19A = data.frame(Plate = ar_plate_nums_19A[1], Plate_well = paste(ar_plate_nums_19A[1], read.table(paste("/Users/epariset/Desktop/Processed-all/BNL19A-ProcessedFiles-Threshold3000/average_well_P", ar_plate_nums_19A[1],'.txt', sep=""), sep="\t", header = TRUE)$Well, sep = "_"), read.table(paste("/Users/epariset/Desktop/Processed-all/BNL19A-ProcessedFiles-Threshold3000/average_well_P", ar_plate_nums_19A[1],'.txt', sep=""), sep="\t", header = TRUE))

ar_table_19A=foci_ar_first_19A

foci_fe_first_19A = data.frame(Plate = fe_plate_nums_19A[1], Plate_well = paste(fe_plate_nums_19A[1], read.table(paste("/Users/epariset/Desktop/Processed-all/BNL19A-ProcessedFiles-Threshold3000/average_well_P", fe_plate_nums_19A[1],'.txt', sep=""), sep="\t", header = TRUE)$Well, sep = "_"), read.table(paste("/Users/epariset/Desktop/Processed-all/BNL19A-ProcessedFiles-Threshold3000/average_well_P", fe_plate_nums_19A[1],'.txt', sep=""), sep="\t", header = TRUE))

fe_table_19A=foci_fe_first_19A
```

```{r}
## Read all the other files and combine them into a single giant file with ALL the plates and ALL the wells

## BNL18B:
for(ii in 2:si_plate_count_18B){
  temp_si_filename = paste('/Users/epariset/Desktop/Processed-all/BNL18B-ProcessedFile/Si-BNL_18B/average_well_P', si_plate_nums_18B[ii],'.txt', sep="")
  temp_si_table_18B = data.frame(Plate = si_plate_nums_18B[ii], Plate_well = paste(si_plate_nums_18B[ii], read.table(paste('/Users/epariset/Desktop/Processed-all/BNL18B-ProcessedFile/Si-BNL_18B/average_well_P', si_plate_nums_18B[ii],'.txt', sep=""), sep="\t", header = TRUE)$Well, sep = "_"), read.table(paste('/Users/epariset/Desktop/Processed-all/BNL18B-ProcessedFile/Si-BNL_18B/average_well_P', si_plate_nums_18B[ii],'.txt', sep=""), sep="\t", header = TRUE))
  si_table_18B = rbind(si_table_18B, temp_si_table_18B)
}

for(ii in 2:ar_plate_count_18B){
  temp_ar_filename = paste('/Users/epariset/Desktop/Processed-all/BNL18B-ProcessedFile/Ar-BNL_18B/average_well_P', ar_plate_nums_18B[ii],'.txt', sep="")
  temp_ar_table_18B = data.frame(Plate = ar_plate_nums_18B[ii], Plate_well = paste(ar_plate_nums_18B[ii], read.table(paste('/Users/epariset/Desktop/Processed-all/BNL18B-ProcessedFile/Ar-BNL_18B/average_well_P', ar_plate_nums_18B[ii],'.txt', sep=""), sep="\t", header = TRUE)$Well, sep = "_"), read.table(paste('/Users/epariset/Desktop/Processed-all/BNL18B-ProcessedFile/Ar-BNL_18B/average_well_P', ar_plate_nums_18B[ii],'.txt', sep=""), sep="\t", header = TRUE))
  ar_table_18B = rbind(ar_table_18B, temp_ar_table_18B)
  }

for(ii in 2:fe_plate_count_18B){
  temp_fe_filename = paste('/Users/epariset/Desktop/Processed-all/BNL19A-ProcessedFiles-Threshold3000/average_well_P', fe_plate_nums_18B[ii],'.txt', sep="")
  temp_fe_table_18B = data.frame(Plate = fe_plate_nums_18B[ii], Plate_well = paste(fe_plate_nums_18B[ii], read.table(paste('/Users/epariset/Desktop/Processed-all/BNL18B-ProcessedFile/Fe-BNL_18B/average_well_P', fe_plate_nums_18B[ii],'.txt', sep=""), sep="\t", header = TRUE)$Well, sep = "_"), read.table(paste('/Users/epariset/Desktop/Processed-all/BNL18B-ProcessedFile/Fe-BNL_18B/average_well_P', fe_plate_nums_18B[ii],'.txt', sep=""), sep="\t", header = TRUE))
  fe_table_18B = rbind(fe_table_18B, temp_fe_table_18B)
  }

## BNL19A:
for(ii in 2:gamma_plate_count_19A){
  temp_gamma_filename = paste('/Users/epariset/Desktop/Processed-all/BNL19A-ProcessedFiles-Threshold3000/average_well_P', gamma_plate_nums_19A[ii],'.txt', sep="")
  temp_gamma_table_19A = data.frame(Plate = gamma_plate_nums_19A[ii], Plate_well = paste(gamma_plate_nums_19A[ii], read.table(paste('/Users/epariset/Desktop/Processed-all/BNL19A-ProcessedFiles-Threshold3000/average_well_P', gamma_plate_nums_19A[ii],'.txt', sep=""), sep="\t", header = TRUE)$Well, sep = "_"), read.table(paste('/Users/epariset/Desktop/Processed-all/BNL19A-ProcessedFiles-Threshold3000/average_well_P', gamma_plate_nums_19A[ii],'.txt', sep=""), sep="\t", header = TRUE))
  gamma_table_19A = rbind(gamma_table_19A, temp_gamma_table_19A)
}

for(ii in 2:si_plate_count_19A){
  temp_si_filename = paste('/Users/epariset/Desktop/Processed-all/BNL19A-ProcessedFiles-Threshold3000/average_well_P', si_plate_nums_19A[ii],'.txt', sep="")
  temp_si_table_19A = data.frame(Plate = si_plate_nums_19A[ii], Plate_well = paste(si_plate_nums_19A[ii], read.table(paste('/Users/epariset/Desktop/Processed-all/BNL19A-ProcessedFiles-Threshold3000/average_well_P', si_plate_nums_19A[ii],'.txt', sep=""), sep="\t", header = TRUE)$Well, sep = "_"), read.table(paste('/Users/epariset/Desktop/Processed-all/BNL19A-ProcessedFiles-Threshold3000/average_well_P', si_plate_nums_19A[ii],'.txt', sep=""), sep="\t", header = TRUE))
  si_table_19A = rbind(si_table_19A, temp_si_table_19A)
}

for(ii in 2:ar_plate_count_19A){
  temp_ar_filename = paste('/Users/epariset/Desktop/Processed-all/BNL19A-ProcessedFiles-Threshold3000/average_well_P', ar_plate_nums_19A[ii],'.txt', sep="")
  temp_ar_table_19A = data.frame(Plate = ar_plate_nums_19A[ii], Plate_well = paste(ar_plate_nums_19A[ii], read.table(paste('/Users/epariset/Desktop/Processed-all/BNL19A-ProcessedFiles-Threshold3000/average_well_P', ar_plate_nums_19A[ii],'.txt', sep=""), sep="\t", header = TRUE)$Well, sep = "_"), read.table(paste('/Users/epariset/Desktop/Processed-all/BNL19A-ProcessedFiles-Threshold3000/average_well_P', ar_plate_nums_19A[ii],'.txt', sep=""), sep="\t", header = TRUE))
  ar_table_19A = rbind(ar_table_19A, temp_ar_table_19A)
  }

for(ii in 2:fe_plate_count_19A){
  temp_fe_filename = paste('/Users/epariset/Desktop/Processed-all/BNL19A-ProcessedFiles-Threshold3000/average_well_P', fe_plate_nums_19A[ii],'.txt', sep="")
  temp_fe_table_19A = data.frame(Plate = fe_plate_nums_19A[ii], Plate_well = paste(fe_plate_nums_19A[ii], read.table(paste('/Users/epariset/Desktop/Processed-all/BNL19A-ProcessedFiles-Threshold3000/average_well_P', fe_plate_nums_19A[ii],'.txt', sep=""), sep="\t", header = TRUE)$Well, sep = "_"), read.table(paste('/Users/epariset/Desktop/Processed-all/BNL19A-ProcessedFiles-Threshold3000/average_well_P', fe_plate_nums_19A[ii],'.txt', sep=""), sep="\t", header = TRUE))
  fe_table_19A = rbind(fe_table_19A, temp_fe_table_19A)
  }
```

```{r}
## Combine runs 18B and 19A
si_table_18B$BNL_run = '18B'
ar_table_18B$BNL_run = '18B'
fe_table_18B$BNL_run = '18B'
si_table_19A$BNL_run = '19A'
ar_table_19A$BNL_run = '19A'
fe_table_19A$BNL_run = '19A'
gamma_table_19A$BNL_run = '19A'

si_table = rbind(si_table_18B,si_table_19A)
ar_table = rbind(ar_table_18B,ar_table_19A)
fe_table = rbind(fe_table_18B,fe_table_19A)
gamma_table = gamma_table_19A
```

```{r}
## Now we will add plate metadata

fe_table_w_plate_meta = merge(fe_table, plate_meta, by.x = 'Plate', by.y = 'Plate_number', all.x = TRUE, all.y = FALSE)
ar_table_w_plate_meta = merge(ar_table, plate_meta, by.x = 'Plate', by.y = 'Plate_number', all.x = TRUE, all.y = FALSE)
si_table_w_plate_meta = merge(si_table, plate_meta, by.x = 'Plate', by.y = 'Plate_number', all.x = TRUE, all.y = FALSE)
gamma_table_w_plate_meta = merge(gamma_table, plate_meta, by.x = 'Plate', by.y = 'Plate_number', all.x = TRUE, all.y = FALSE)

## Now we will add well metadata

# Wells are different based on sample sets, so we need to make an extra column for Set_Well, and then add well metadata based on it.
fe_table_w_plate_meta$Set_well = paste(fe_table_w_plate_meta$Sample_sets, fe_table_w_plate_meta$Well, sep = "_")
subject_meta$Set_well = paste(subject_meta$Sample_ID, subject_meta$Well, sep = "_")
fe_table_w_meta = merge(fe_table_w_plate_meta, subject_meta, by.x = 'Set_well', by.y = 'Set_Well')

ar_table_w_plate_meta$Set_well = paste(ar_table_w_plate_meta$Sample_sets, ar_table_w_plate_meta$Well, sep = "_")
subject_meta$Set_well = paste(subject_meta$Sample_ID, subject_meta$Well, sep = "_")
ar_table_w_meta = merge(ar_table_w_plate_meta, subject_meta, by.x = 'Set_well', by.y = 'Set_Well')

si_table_w_plate_meta$Set_well = paste(si_table_w_plate_meta$Sample_sets, si_table_w_plate_meta$Well, sep = "_")
subject_meta$Set_well = paste(subject_meta$Sample_ID, subject_meta$Well, sep = "_")
si_table_w_meta = merge(si_table_w_plate_meta, subject_meta, by.x ="Set_well", by.y="Set_Well", all.x=TRUE)

gamma_table_w_plate_meta$Set_well = paste(gamma_table_w_plate_meta$Sample_sets, gamma_table_w_plate_meta$Well, sep = "_")
subject_meta$Set_well = paste(subject_meta$Sample_ID, subject_meta$Well, sep = "_")
gamma_table_w_meta = merge(gamma_table_w_plate_meta, subject_meta, by.x ="Set_well", by.y="Set_Well", all.x=TRUE)

# Clean up: this has Well.x and Well.y, which are duplicates; Sample_sets and Set which are also duplicates.
fe_table_w_meta$Well = fe_table_w_meta$Well.x
fe_table_w_meta$Timepoint = fe_table_w_meta$Timepoint.x
cols_to_drop = c('Well.x', 'Well.y', 'Sample_sets', 'Plate.1')
fe_long = fe_table_w_meta[, !colnames(fe_table_w_meta) %in% cols_to_drop]

ar_table_w_meta$Well = ar_table_w_meta$Well.x 
ar_table_w_meta$Timepoint = ar_table_w_meta$Timepoint.x
cols_to_drop = c('Well.x', 'Well.y', 'Sample_sets', 'Timepoint.x', 'Timepoint.y')
ar_long = ar_table_w_meta[, !colnames(ar_table_w_meta) %in% cols_to_drop]

si_table_w_meta$Well = si_table_w_meta$Well.x
si_table_w_meta$Timepoint = si_table_w_meta$Timepoint.x
cols_to_drop = c('Well.x', 'Well.y', 'Sample_sets', 'Timepoint.x', 'Timepoint.y')
si_long = si_table_w_meta[, !colnames(si_table_w_meta) %in% cols_to_drop]

gamma_table_w_meta$Well = gamma_table_w_meta$Well.x
gamma_table_w_meta$Timepoint = gamma_table_w_meta$Timepoint.x
cols_to_drop = c('Well.x', 'Well.y', 'Sample_sets', 'Timepoint.x', 'Timepoint.y')
gamma_long = gamma_table_w_meta[, !colnames(gamma_table_w_meta) %in% cols_to_drop]
```

```{r}
# Set low bar. This is arbitrary! We will ask for at least 50 imaged nuclei per well to consider that well. 
low_bar = 50
fe_filt = fe_long[fe_long$num_nuc > (low_bar-1),]
ar_filt = ar_long[ar_long$num_nuc > (low_bar-1),]
si_filt = si_long[si_long$num_nuc > (low_bar-1),]
gamma_filt = gamma_long[gamma_long$num_nuc > (low_bar-1),]

print("How many wells are left after filtering for Fe?")
print(c(nrow(fe_filt),"out of", nrow(fe_long), "which is", (nrow(fe_filt)/nrow(fe_long))*100,"%"))

print("How many wells are left after filtering for Ar?")
print(c(nrow(ar_filt),"out of", nrow(ar_long), "which is", (nrow(ar_filt)/nrow(ar_long))*100,"%"))

print("How many wells are left after filtering for Si?")
print(c(nrow(si_filt),"out of", nrow(si_long), "which is", (nrow(si_filt)/nrow(si_long))*100,"%"))

print("How many wells are left after filtering for Gamma?")
print(c(nrow(gamma_filt),"out of", nrow(gamma_long), "which is", (nrow(gamma_filt)/nrow(gamma_long))*100,"%"))
```

```{r}
# Truncate dataframe to only include the columns of interest
col_trunc = c("Sample_ID", "Duplicates", "avg_nfoci", "num_nuc", "Set", "Rad_type", "Dose_Fluence", "Timepoint", "Age", "Gender", "BMI", "BMI_group", "baseline_avg_nfoci", "BNL_run")
fe_trunc = fe_filt[,colnames(fe_filt) %in% col_trunc]
fe_trunc$ID_Fluence_Timepoint = paste(fe_trunc$Sample_ID, fe_trunc$Dose_Fluence, fe_trunc$Timepoint, sep = "_")

col_trunc = c("Sample_ID", "Duplicates", "avg_nfoci", "num_nuc", "Set", "Rad_type", "Dose_Fluence", "Timepoint", "Age", "Gender", "BMI", "BMI_group", "baseline_avg_nfoci", "BNL_run")
ar_trunc = ar_filt[,colnames(ar_filt) %in% col_trunc]
ar_trunc$ID_Fluence_Timepoint = paste(ar_trunc$Sample_ID, ar_trunc$Dose_Fluence, ar_trunc$Timepoint, sep = "_")

col_trunc = c("Sample_ID", "Duplicates", "avg_nfoci", "num_nuc", "Set", "Rad_type", "Dose_Fluence", "Timepoint", "Age", "Gender", "BMI", "BMI_group", "baseline_avg_nfoci", "BNL_run")
si_trunc = si_filt[,colnames(si_filt) %in% col_trunc]
si_trunc$ID_Fluence_Timepoint = paste(si_trunc$Sample_ID, si_trunc$Dose_Fluence, si_trunc$Timepoint, sep = "_")

col_trunc = c("Sample_ID", "Duplicates", "avg_nfoci", "num_nuc", "Set", "Rad_type", "Dose_Fluence", "Timepoint", "Age", "Gender", "BMI", "BMI_group", "baseline_avg_nfoci", "BNL_run")
gamma_trunc = gamma_filt[,colnames(gamma_filt) %in% col_trunc]
gamma_trunc$ID_Fluence_Timepoint = paste(gamma_trunc$Sample_ID, gamma_trunc$Dose_Fluence, gamma_trunc$Timepoint, sep = "_")
```

```{r}
# Now we will need to average A and B duplicates
fe_avg = ddply(fe_trunc, .(ID_Fluence_Timepoint, Sample_ID, Set, Rad_type, Dose_Fluence, Timepoint, Age, Gender, BMI, BMI_group, baseline_avg_nfoci, BNL_run), summarise, avg_nfoci = mean(avg_nfoci), num_nuc = mean(num_nuc))
ar_avg = ddply(ar_trunc, .(ID_Fluence_Timepoint, Sample_ID, Set, Rad_type, Dose_Fluence, Timepoint, Age, Gender, BMI, BMI_group, baseline_avg_nfoci, BNL_run), summarise, num_nuc = mean(num_nuc), avg_nfoci = mean(avg_nfoci))
si_avg = ddply(si_trunc, .(ID_Fluence_Timepoint, Sample_ID, Set, Rad_type, Dose_Fluence, Timepoint, Age, Gender, BMI, BMI_group, baseline_avg_nfoci, BNL_run), summarise, num_nuc = mean(num_nuc), avg_nfoci = mean(avg_nfoci))
gamma_avg = ddply(gamma_trunc, .(ID_Fluence_Timepoint, Sample_ID, Set, Rad_type, Dose_Fluence, Timepoint, Age, Gender, BMI, BMI_group, baseline_avg_nfoci, BNL_run), summarise, avg_nfoci = mean(avg_nfoci), num_nuc = mean(num_nuc))
```

```{r}
# I don't like calling "Dose_Fluence" when we only have fluence or dose; also it is a factor instead of a number
colnames(fe_avg)[colnames(fe_avg) == "Dose_Fluence"] = "Fluence"
fe_avg$Fluence = as.factor(fe_avg$Fluence)

colnames(ar_avg)[colnames(ar_avg) == "Dose_Fluence"] = "Fluence"
ar_avg$Fluence = as.factor(ar_avg$Fluence)

colnames(si_avg)[colnames(si_avg) == "Dose_Fluence"] = "Fluence"
si_avg$Fluence = as.factor(si_avg$Fluence)

colnames(gamma_avg)[colnames(gamma_avg) == "Dose_Fluence"] = "Dose"
gamma_avg$Dose = as.factor(gamma_avg$Dose)
```

```{r}
# Also I want my timepoint to go from 4h to 24h in outputs
fe_avg$Timepoint = factor(fe_avg$Timepoint, levels = c("4h", "24h")) 
ar_avg$Timepoint = factor(ar_avg$Timepoint, levels = c("4h", "24h")) 
si_avg$Timepoint = factor(si_avg$Timepoint, levels = c("4h", "24h")) 
gamma_avg$Timepoint = factor(gamma_avg$Timepoint, levels = c("4h", "24h")) 
```

```{r}
# Also I want my BMI group to go from Underweight to Normal to Overweight to Obese in outputs
fe_avg$BMI_group = factor(fe_avg$BMI_group, levels = c("Underweight", "Normal", "Overweight", "Obese")) 
ar_avg$BMI_group = factor(ar_avg$BMI_group, levels = c("Underweight", "Normal", "Overweight", "Obese"))
si_avg$BMI_group = factor(si_avg$BMI_group, levels = c("Underweight", "Normal", "Overweight", "Obese")) 
gamma_avg$BMI_group = factor(gamma_avg$BMI_group, levels = c("Underweight", "Normal", "Overweight", "Obese"))
```

-------------------------------------------------------------------------------------------------------------------------------------------------------------

```{r}
# First, let's see how avg_nfoci relates to the fluence/dose for each irradiation at 4h and 24h (as a boxplot form)
all_plot = ggplot(fe_avg, aes(x = Fluence, y = avg_nfoci, col = Gender)) + 
geom_boxplot() + ggtitle("Iron")
facet(all_plot, facet.by = c('Timepoint', 'BNL_run'))

all_plot = ggplot(ar_avg, aes(x = Fluence, y = avg_nfoci, col = Gender)) + 
geom_boxplot() + ggtitle("Argon")
facet(all_plot, facet.by = 'Timepoint')

all_plot = ggplot(si_avg, aes(x = Fluence, y = avg_nfoci, col = Gender)) + 
geom_boxplot() + ggtitle("Silicon")
facet(all_plot, facet.by = 'Timepoint')

all_plot = ggplot(gamma_avg, aes(x = Dose, y = avg_nfoci, col = Gender)) + 
geom_boxplot() + ggtitle("Gamma")
facet(all_plot, facet.by = 'Timepoint')
```

```{r}
# Then, let's see how avg_nfoci relates to the fluence/dose for each irradiation at 4h and 24h as individual samples

fe_avg$Fluence_num = as.numeric(as.character(fe_avg$Fluence)) # This is to add slopes of avg_nfoci = f(fluence)
fe_avg_1st_slope = fe_avg[(fe_avg$Fluence_num != "3"),] # First slope between fluences 0 and 1.1
fe_avg_2nd_slope = fe_avg[(fe_avg$Fluence_num != "0"),] # Second slope between fluence 1.1 and 3

all_plot_dots_fe_slopes = ggplot(fe_avg, aes(x = Fluence_num, y = avg_nfoci, col = Gender)) + geom_point() + geom_jitter() + 
  geom_smooth(data = fe_avg_1st_slope, method = "glm") +
  geom_smooth(data = fe_avg_2nd_slope, method = "glm") + ggtitle("Iron") + xlab("Fluence")
facet(all_plot_dots_fe_slopes, facet.by = 'Timepoint')

ar_avg$Fluence_num = as.numeric(as.character(ar_avg$Fluence)) # This is to add slopes of avg_nfoci = f(fluence)
ar_avg_1st_slope = ar_avg[(ar_avg$Fluence_num != "3"),] # First slope between fluences 0 and 1.1
ar_avg_2nd_slope = ar_avg[(ar_avg$Fluence_num != "0"),] # Second slope between fluence 1.1 and 3

all_plot_dots_ar_slopes = ggplot(ar_avg, aes(x = Fluence_num, y = avg_nfoci, col = Gender)) + geom_point() + geom_jitter() + 
  geom_smooth(data = ar_avg_1st_slope, method = "glm") +
  geom_smooth(data = ar_avg_2nd_slope, method = "glm") + ggtitle("Argon") + xlab("Fluence")
facet(all_plot_dots_ar_slopes, facet.by = 'Timepoint')

si_avg$Fluence_num = as.numeric(as.character(si_avg$Fluence)) # This is to add slopes of avg_nfoci = f(fluence)
si_avg_1st_slope = si_avg[(si_avg$Fluence_num != "3"),] # First slope between fluences 0 and 1.1
si_avg_2nd_slope = si_avg[(si_avg$Fluence_num != "0"),] # Second slope between fluence 1.1 and 3

all_plot_dots_si_slopes = ggplot(si_avg, aes(x = Fluence_num, y = avg_nfoci, col = Gender)) + geom_point() + geom_jitter() + 
  geom_smooth(data = si_avg_1st_slope, method = "glm") +
  geom_smooth(data = si_avg_2nd_slope, method = "glm") + ggtitle("Silicon") + xlab("Fluence")
facet(all_plot_dots_si_slopes, facet.by = 'Timepoint')

gamma_avg$Dose_num = as.numeric(as.character(gamma_avg$Dose)) # This is to add slopes of avg_nfoci = f(dose)
gamma_avg_1st_slope = gamma_avg[(gamma_avg$Dose_num != "1"),] # First slope between fluences 0 and 0.1
gamma_avg_2nd_slope = gamma_avg[(gamma_avg$Dose_num != "0"),] # Second slope between fluence 0.1 and 1

all_plot_dots_gamma_slopes = ggplot(gamma_avg, aes(x = Dose_num, y = avg_nfoci, col = Gender)) + geom_point() + geom_jitter() + 
  geom_smooth(data = gamma_avg_1st_slope, method = "glm") +
  geom_smooth(data = gamma_avg_2nd_slope, method = "glm") + ggtitle("Gamma")
facet(all_plot_dots_gamma_slopes, facet.by = 'Timepoint')
```

```{r}
# We want to check that each individual has similar avg_nfoci in all 4 radiation types at 4h and 24h
sham_fe_avg_4h = fe_avg[fe_avg$Fluence == "0" & fe_avg$Timepoint == "4h",]
sham_ar_avg_4h = ar_avg[ar_avg$Fluence == "0" & ar_avg$Timepoint == "4h",]
sham_si_avg_4h = si_avg[si_avg$Fluence == "0" & si_avg$Timepoint == "4h",]
sham_gamma_avg_4h = gamma_avg[gamma_avg$Fluence == "0" & gamma_avg$Timepoint == "4h",]
all_sham_avg_4h = rbind(sham_fe_avg_4h, sham_ar_avg_4h, sham_si_avg_4h, sham_gamma_avg_4h)

all_plot = ggplot(all_sham_avg_4h, aes(x = Sample_ID, y = avg_nfoci, col = Rad_type)) + 
  geom_point()
all_plot

sham_fe_avg_24h = fe_avg[fe_avg$Fluence == "0" & fe_avg$Timepoint == "24h",]
sham_ar_avg_24h = ar_avg[ar_avg$Fluence == "0" & ar_avg$Timepoint == "24h",]
sham_si_avg_24h = si_avg[si_avg$Fluence == "0" & si_avg$Timepoint == "24h",]
sham_gamma_avg_24h = gamma_avg[gamma_avg$Dose == "0" & gamma_avg$Timepoint == "24h",]
colnames(sham_gamma_avg_24h)[5] <- "Fluence"
all_sham_avg_24h = rbind(sham_fe_avg_24h, sham_ar_avg_24h, sham_si_avg_24h, sham_gamma_avg_24h)

all_plot = ggplot(all_sham_avg_24h, aes(x = Sample_ID, y = avg_nfoci, col = Rad_type)) + 
  geom_point()
all_plot
```

```{r}
# Now we combine all data for all 4 irradiation types and plot as a function of dose

fe_avg_0 = fe_avg[fe_avg$Fluence == "0",]
fe_avg_0$dose = "0"
fe_avg_dose1 = fe_avg[fe_avg$Fluence == "1.1",]
fe_avg_dose1$dose = "0.31"
fe_avg_dose2 = fe_avg[fe_avg$Fluence == "3",]
fe_avg_dose2$dose = "0.83"

ar_avg_0 = ar_avg[ar_avg$Fluence == "0",]
ar_avg_0$dose = "0"
ar_avg_dose1 = ar_avg[ar_avg$Fluence == "1.1",]
ar_avg_dose1$dose = "0.18"
ar_avg_dose2 = ar_avg[ar_avg$Fluence == "3",]
ar_avg_dose2$dose = "0.5"

si_avg_0 = si_avg[si_avg$Fluence == "0",]
si_avg_0$dose = "0"
si_avg_dose1 = si_avg[si_avg$Fluence == "1.1",]
si_avg_dose1$dose = "0.11"
si_avg_dose2 = si_avg[si_avg$Fluence == "3",]
si_avg_dose2$dose = "0.29"

colnames(gamma_avg)[5] <- "Fluence"
colnames(gamma_avg)[14] <- "Fluence_num"
gamma_avg_0 = gamma_avg[gamma_avg$Fluence == "0",]
gamma_avg_0$dose = "0"
gamma_avg_dose1 = gamma_avg[gamma_avg$Fluence == "0.1",]
gamma_avg_dose1$dose = "0.1"
gamma_avg_dose2 = gamma_avg[gamma_avg$Fluence == "1",]
gamma_avg_dose2$dose = "1"

all_dose = rbind(fe_avg_0,fe_avg_dose1,fe_avg_dose2,ar_avg_0,ar_avg_dose1,ar_avg_dose2,si_avg_0,si_avg_dose1,si_avg_dose2,gamma_avg_0,gamma_avg_dose1,gamma_avg_dose2)

colnames(gamma_avg)[5] <- "Dose"
colnames(gamma_avg)[14] <- "Dose_num"

dose_plot = ggplot(all_dose, aes(x = dose, y = avg_nfoci, fill = Rad_type)) + 
  geom_boxplot()
facet(dose_plot, facet.by = 'Timepoint')
```

# Study of demographic variables ----------------------------------------------------------------------------------------------------------------------------

```{r}
# Influence of gender and BMI
all_plot_bmi_fe = ggplot(fe_avg, aes(x = Fluence, y = avg_nfoci, col = BMI_group)) + 
  geom_boxplot() + ggtitle("Iron")
facet(all_plot_bmi_fe, facet.by = c('Timepoint', 'Gender'))

all_plot_bmi2_fe = ggplot(fe_avg, aes(x = BMI, y = avg_nfoci, col = Fluence)) + 
  geom_point() + geom_smooth(method = "lm") + ggtitle("Iron")
facet(all_plot_bmi2_fe, facet.by = c('Timepoint', 'Gender'))

all_plot_bmi_ar = ggplot(ar_avg, aes(x = Fluence, y = avg_nfoci, col = BMI_group)) + 
  geom_boxplot() + ggtitle("Argon")
facet(all_plot_bmi_ar, facet.by = c('Timepoint', 'Gender'))

all_plot_bmi2_ar = ggplot(ar_avg, aes(x = BMI, y = avg_nfoci, col = Fluence)) + 
  geom_point() + geom_smooth(method = "lm") + ggtitle("Argon")
facet(all_plot_bmi2_ar, facet.by = c('Timepoint', 'Gender'))

all_plot_bmi_si = ggplot(si_avg, aes(x = Fluence, y = avg_nfoci, col = BMI_group)) + 
  geom_boxplot() + ggtitle("Silicon")
facet(all_plot_bmi_si, facet.by = c('Timepoint', 'Gender'))

all_plot_bmi2_si = ggplot(si_avg, aes(x = BMI, y = avg_nfoci, col = Fluence)) + 
  geom_point() + geom_smooth(method = "lm") + ggtitle("Silicon")
facet(all_plot_bmi2_si, facet.by = c('Timepoint', 'Gender'))

all_plot_bmi_gamma = ggplot(gamma_avg, aes(x = Dose, y = avg_nfoci, col = BMI_group)) + 
  geom_boxplot() + ggtitle("Gamma")
facet(all_plot_bmi_gamma, facet.by = c('Timepoint', 'Gender'))

all_plot_bmi2_gamma = ggplot(gamma_avg, aes(x = BMI, y = avg_nfoci, col = Dose)) + 
  geom_point() + geom_smooth(method = "lm") + ggtitle("Gamma") 
facet(all_plot_bmi2_gamma, facet.by = c('Timepoint', 'Gender'))
```

```{r}
# Influence of age 
all_plot_age_fe = ggplot(fe_avg, aes(x = Age, y = avg_nfoci, col = Fluence)) + 
  geom_point() + geom_smooth(method = "lm") + ggtitle("Iron")
facet(all_plot_age_fe, facet.by = c('Timepoint', 'Gender'))

all_plot_age_ar = ggplot(ar_avg, aes(x = Age, y = avg_nfoci, col = Fluence)) + 
  geom_point() + geom_smooth(method = "lm") + ggtitle("Argon")
facet(all_plot_age_ar, facet.by = c('Timepoint', 'Gender'))

all_plot_age_si = ggplot(si_avg, aes(x = Age, y = avg_nfoci, col = Fluence)) + 
  geom_point() + geom_smooth(method = "lm") + ggtitle("Silicon")
facet(all_plot_age_si, facet.by = c('Timepoint', 'Gender'))

all_plot_age_gamma = ggplot(gamma_avg, aes(x = Age, y = avg_nfoci, col = Dose)) + 
  geom_point() + geom_smooth(method = "lm") + ggtitle("Gamma")
facet(all_plot_age_gamma, facet.by = c('Timepoint', 'Gender'))
```

# Normalization----------------------------------------------------------------------------------------------------------------------------------------------

```{r}
# We want to normalize both fluences/doses to the sham-irradiated control

# We'll have one table at 4h and one table at 24h
fe_avg_4h = fe_avg[fe_avg$Timepoint == "4h",]
fe_avg_24h = fe_avg[fe_avg$Timepoint == "24h",]
ar_avg_4h = ar_avg[ar_avg$Timepoint == "4h",]
ar_avg_24h = ar_avg[ar_avg$Timepoint == "24h",]
si_avg_4h = si_avg[si_avg$Timepoint == "4h",]
si_avg_24h = si_avg[si_avg$Timepoint == "24h",]
gamma_avg_4h = gamma_avg[gamma_avg$Timepoint == "4h",]
gamma_avg_24h = gamma_avg[gamma_avg$Timepoint == "24h",]

fe_avg_4h_norm <- dcast(fe_avg_4h, Sample_ID ~ Fluence, value.var = 'avg_nfoci') #reshape the table fe_avg_4h
fe_avg_24h_norm <- dcast(fe_avg_24h, Sample_ID ~ Fluence, value.var = 'avg_nfoci') #reshape the table fe_avg_24h
ar_avg_4h_norm <- dcast(ar_avg_4h, Sample_ID ~ Fluence, value.var = 'avg_nfoci') #reshape the table ar_avg_4h
ar_avg_24h_norm <- dcast(ar_avg_24h, Sample_ID ~ Fluence, value.var = 'avg_nfoci') #reshape the table ar_avg_24h
si_avg_4h_norm <- dcast(si_avg_4h, Sample_ID ~ Fluence, value.var = 'avg_nfoci') #reshape the table si_avg_4h
si_avg_24h_norm <- dcast(si_avg_24h, Sample_ID ~ Fluence, value.var = 'avg_nfoci') #reshape the table si_avg_24h
gamma_avg_4h_norm <- dcast(gamma_avg_4h, Sample_ID ~ Dose, value.var = 'avg_nfoci') #reshape the table gamma_avg_4h
gamma_avg_24h_norm <- dcast(gamma_avg_24h, Sample_ID ~ Dose, value.var = 'avg_nfoci') #reshape the table gamma_avg_24h
```

```{r}
# We choose to keep only the data for which avg_nfoci at dose 1 and dose 2 >= avg_nfoci at dose 0
fe_avg_4h_norm = fe_avg_4h_norm[fe_avg_4h_norm$'1.1' >= fe_avg_4h_norm$'0',]
fe_avg_4h_norm = fe_avg_4h_norm[fe_avg_4h_norm$'3' >= fe_avg_4h_norm$'0',]
fe_avg_4h_norm <- na.omit(fe_avg_4h_norm)
fe_avg_24h_norm = fe_avg_24h_norm[fe_avg_24h_norm$'1.1' >= fe_avg_24h_norm$'0',]
fe_avg_24h_norm = fe_avg_24h_norm[fe_avg_24h_norm$'3' >= fe_avg_24h_norm$'0',]
fe_avg_24h_norm <- na.omit(fe_avg_24h_norm)

ar_avg_4h_norm = ar_avg_4h_norm[ar_avg_4h_norm$'1.1' >= ar_avg_4h_norm$'0',]
ar_avg_4h_norm = ar_avg_4h_norm[ar_avg_4h_norm$'3' >= ar_avg_4h_norm$'0',]
ar_avg_4h_norm <- na.omit(ar_avg_4h_norm)
ar_avg_24h_norm = ar_avg_24h_norm[ar_avg_24h_norm$'1.1' >= ar_avg_24h_norm$'0',]
ar_avg_24h_norm = ar_avg_24h_norm[ar_avg_24h_norm$'3' >= ar_avg_24h_norm$'0',]
ar_avg_24h_norm <- na.omit(ar_avg_24h_norm)

si_avg_4h_norm = si_avg_4h_norm[si_avg_4h_norm$'1.1' >= si_avg_4h_norm$'0',]
si_avg_4h_norm = si_avg_4h_norm[si_avg_4h_norm$'3' >= si_avg_4h_norm$'0',]
si_avg_4h_norm <- na.omit(si_avg_4h_norm)
si_avg_24h_norm = si_avg_24h_norm[si_avg_24h_norm$'1.1' >= si_avg_24h_norm$'0',]
si_avg_24h_norm = si_avg_24h_norm[si_avg_24h_norm$'3' >= si_avg_24h_norm$'0',]
si_avg_24h_norm <- na.omit(si_avg_24h_norm)

gamma_avg_4h_norm = gamma_avg_4h_norm[gamma_avg_4h_norm$'0.1' >= gamma_avg_4h_norm$'0',]
gamma_avg_4h_norm = gamma_avg_4h_norm[gamma_avg_4h_norm$'1' >= gamma_avg_4h_norm$'0',]
gamma_avg_4h_norm <- na.omit(gamma_avg_4h_norm)
gamma_avg_24h_norm = gamma_avg_24h_norm[gamma_avg_24h_norm$'0.1' >= gamma_avg_24h_norm$'0',]
gamma_avg_24h_norm = gamma_avg_24h_norm[gamma_avg_24h_norm$'1' >= gamma_avg_24h_norm$'0',]
gamma_avg_24h_norm <- na.omit(gamma_avg_24h_norm)
```

```{r}
# We normalize by dividing avg_nfoci for each of the doses to the sham-irradiated control
fe_avg_4h_norm$norm_nfoci1 = fe_avg_4h_norm$'1.1'/fe_avg_4h_norm$'0'
fe_avg_4h_norm$norm_nfoci2 = fe_avg_4h_norm$'3'/fe_avg_4h_norm$'0'
fe_avg_24h_norm$norm_nfoci1 = fe_avg_24h_norm$'1.1'/fe_avg_24h_norm$'0'
fe_avg_24h_norm$norm_nfoci2 = fe_avg_24h_norm$'3'/fe_avg_24h_norm$'0'

ar_avg_4h_norm$norm_nfoci1 = ar_avg_4h_norm$'1.1'/ar_avg_4h_norm$'0'
ar_avg_4h_norm$norm_nfoci2 = ar_avg_4h_norm$'3'/ar_avg_4h_norm$'0'
ar_avg_24h_norm$norm_nfoci1 = ar_avg_24h_norm$'1.1'/ar_avg_24h_norm$'0'
ar_avg_24h_norm$norm_nfoci2 = ar_avg_24h_norm$'3'/ar_avg_24h_norm$'0'

si_avg_4h_norm$norm_nfoci1 = si_avg_4h_norm$'1.1'/si_avg_4h_norm$'0'
si_avg_4h_norm$norm_nfoci2 = si_avg_4h_norm$'3'/si_avg_4h_norm$'0'
si_avg_24h_norm$norm_nfoci1 = si_avg_24h_norm$'1.1'/si_avg_24h_norm$'0'
si_avg_24h_norm$norm_nfoci2 = si_avg_24h_norm$'3'/si_avg_24h_norm$'0'

gamma_avg_4h_norm$norm_nfoci1 = gamma_avg_4h_norm$'0.1'/gamma_avg_4h_norm$'0'
gamma_avg_4h_norm$norm_nfoci2 = gamma_avg_4h_norm$'1'/gamma_avg_4h_norm$'0'
gamma_avg_24h_norm$norm_nfoci1 = gamma_avg_24h_norm$'0.1'/gamma_avg_24h_norm$'0'
gamma_avg_24h_norm$norm_nfoci2 = gamma_avg_24h_norm$'1'/gamma_avg_24h_norm$'0'
```

```{r}
# We reshape the table to keep only normalized data and combine all radiation types
col_trunc = c("Sample_ID","norm_nfoci1", "norm_nfoci2")
fe_avg_4h_norm = fe_avg_4h_norm[,colnames(fe_avg_4h_norm) %in% col_trunc]
colnames(fe_avg_4h_norm)[colnames(fe_avg_4h_norm) == "norm_nfoci1"] = "1.1"
colnames(fe_avg_4h_norm)[colnames(fe_avg_4h_norm) == "norm_nfoci2"] = "3"
fe_avg_4h_norm <- melt(fe_avg_4h_norm, id = c("Sample_ID"))
fe_avg_4h_norm$Rad_type = "Fe"
fe_avg_24h_norm = fe_avg_24h_norm[,colnames(fe_avg_24h_norm) %in% col_trunc]
colnames(fe_avg_24h_norm)[colnames(fe_avg_24h_norm) == "norm_nfoci1"] = "1.1"
colnames(fe_avg_24h_norm)[colnames(fe_avg_24h_norm) == "norm_nfoci2"] = "3"
fe_avg_24h_norm <- melt(fe_avg_24h_norm, id = c("Sample_ID"))
fe_avg_24h_norm$Rad_type = "Fe"

ar_avg_4h_norm = ar_avg_4h_norm[,colnames(ar_avg_4h_norm) %in% col_trunc]
colnames(ar_avg_4h_norm)[colnames(ar_avg_4h_norm) == "norm_nfoci1"] = "1.1"
colnames(ar_avg_4h_norm)[colnames(ar_avg_4h_norm) == "norm_nfoci2"] = "3"
ar_avg_4h_norm <- melt(ar_avg_4h_norm, id = c("Sample_ID"))
ar_avg_4h_norm$Rad_type = "Ar"
ar_avg_24h_norm = ar_avg_24h_norm[,colnames(ar_avg_24h_norm) %in% col_trunc]
colnames(ar_avg_24h_norm)[colnames(ar_avg_24h_norm) == "norm_nfoci1"] = "1.1"
colnames(ar_avg_24h_norm)[colnames(ar_avg_24h_norm) == "norm_nfoci2"] = "3"
ar_avg_24h_norm <- melt(ar_avg_24h_norm, id = c("Sample_ID"))
ar_avg_24h_norm$Rad_type = "Ar"

si_avg_4h_norm = si_avg_4h_norm[,colnames(si_avg_4h_norm) %in% col_trunc]
colnames(si_avg_4h_norm)[colnames(si_avg_4h_norm) == "norm_nfoci1"] = "1.1"
colnames(si_avg_4h_norm)[colnames(si_avg_4h_norm) == "norm_nfoci2"] = "3"
si_avg_4h_norm <- melt(si_avg_4h_norm, id = c("Sample_ID"))
si_avg_4h_norm$Rad_type = "Si"
si_avg_24h_norm = si_avg_24h_norm[,colnames(si_avg_24h_norm) %in% col_trunc]
colnames(si_avg_24h_norm)[colnames(si_avg_24h_norm) == "norm_nfoci1"] = "1.1"
colnames(si_avg_24h_norm)[colnames(si_avg_24h_norm) == "norm_nfoci2"] = "3"
si_avg_24h_norm <- melt(si_avg_24h_norm, id = c("Sample_ID"))
si_avg_24h_norm$Rad_type = "Si"

gamma_avg_4h_norm = gamma_avg_4h_norm[,colnames(gamma_avg_4h_norm) %in% col_trunc]
colnames(gamma_avg_4h_norm)[colnames(gamma_avg_4h_norm) == "norm_nfoci1"] = "0.1"
colnames(gamma_avg_4h_norm)[colnames(gamma_avg_4h_norm) == "norm_nfoci2"] = "1"
gamma_avg_4h_norm <- melt(gamma_avg_4h_norm, id = c("Sample_ID"))
gamma_avg_4h_norm$Rad_type = "Gamma"
gamma_avg_24h_norm = gamma_avg_24h_norm[,colnames(gamma_avg_24h_norm) %in% col_trunc]
colnames(gamma_avg_24h_norm)[colnames(gamma_avg_24h_norm) == "norm_nfoci1"] = "0.1"
colnames(gamma_avg_24h_norm)[colnames(gamma_avg_24h_norm) == "norm_nfoci2"] = "1"
gamma_avg_24h_norm <- melt(gamma_avg_24h_norm, id = c("Sample_ID"))
gamma_avg_24h_norm$Rad_type = "Gamma"

colnames(gamma_avg_4h_norm)[colnames(gamma_avg_4h_norm) == "variable"] = "Dose"
colnames(gamma_avg_4h_norm)[colnames(gamma_avg_4h_norm) == "value"] = "norm_nfoci"
colnames(gamma_avg_24h_norm)[colnames(gamma_avg_24h_norm) == "variable"] = "Dose"
colnames(gamma_avg_24h_norm)[colnames(gamma_avg_24h_norm) == "value"] = "norm_nfoci"

all_4h_norm <- rbind(fe_avg_4h_norm, ar_avg_4h_norm, si_avg_4h_norm)
colnames(all_4h_norm)[colnames(all_4h_norm) == "variable"] = "Fluence"
colnames(all_4h_norm)[colnames(all_4h_norm) == "value"] = "norm_nfoci"
all_24h_norm <- rbind(fe_avg_24h_norm, ar_avg_24h_norm, si_avg_24h_norm)
colnames(all_24h_norm)[colnames(all_24h_norm) == "variable"] = "Fluence"
colnames(all_24h_norm)[colnames(all_24h_norm) == "value"] = "norm_nfoci"
```

```{r}
# We want the ions to go from Si to Ar to Fe
all_4h_norm$Rad_type = factor(all_4h_norm$Rad_type, levels = c("Si", "Ar", "Fe")) 
all_24h_norm$Rad_type = factor(all_24h_norm$Rad_type, levels = c("Si", "Ar", "Fe"))
```

```{r}
# Now, let's plot it for ions together + gamma separately
all_plot = ggplot(all_4h_norm, aes(x = Fluence, y = norm_nfoci, col = Rad_type)) + 
  geom_boxplot() + ggtitle("Time point: 4h post-irradiation") 
all_plot  + ylim(1,2.5) + xlab("Fluence (particles/100um^2)") + ylab("Normalized Average Number of Foci per Nucleus") + theme(axis.text.x = element_text(size=18)) + theme(axis.text.y = element_text(size=18)) + theme(axis.title.x = element_text(size=18)) + theme(axis.title.y = element_text(size=18)) + labs(col="Radiation") + theme(plot.title = element_text(size=18, face="bold"), legend.text=element_text(size=16), legend.title=element_text(size=16))

all_plot = ggplot(gamma_avg_4h_norm, aes(x = Dose, y = norm_nfoci)) + 
  geom_boxplot() + ggtitle("Time point: 4h post-irradiation")
all_plot + ylim(1,2.5) + xlab("Dose (Gy)") + ylab("Normalized Average Number of Foci per Nucleus") + theme(axis.text.x = element_text(size=18)) + theme(axis.text.y = element_text(size=18)) + theme(axis.title.x = element_text(size=18)) + theme(axis.title.y = element_text(size=18)) + labs(col="Radiation") + theme(plot.title = element_text(size=18, face="bold"), legend.text=element_text(size=16), legend.title=element_text(size=16))

all_plot = ggplot(all_24h_norm, aes(x = Fluence, y = norm_nfoci, col = Rad_type)) + 
  geom_boxplot(lwd=1) + ggtitle("Time point: 24h post-irradiation")
all_plot + ylim(1,2.5) + xlab("Fluence (particles/100um^2)") + ylab("Normalized Average Number of Foci per Nucleus") + theme(axis.text.x = element_text(size=18)) + theme(axis.text.y = element_text(size=18)) + theme(axis.title.x = element_text(size=18)) + theme(axis.title.y = element_text(size=18)) + labs(col="Radiation") + theme(plot.title = element_text(size=18, face="bold"), legend.text=element_text(size=16), legend.title=element_text(size=16))

all_plot = ggplot(gamma_avg_24h_norm, aes(x = Dose, y = norm_nfoci)) + 
  geom_boxplot() + ggtitle("Time point: 24h post-irradiation")
all_plot + ylim(1,2.5) + xlab("Dose (Gy)") + ylab("Normalized Average Number of Foci per Nucleus") + theme(axis.text.x = element_text(size=18)) + theme(axis.text.y = element_text(size=18)) + theme(axis.title.x = element_text(size=18)) + theme(axis.title.y = element_text(size=18)) + labs(col="Radiation") + theme(plot.title = element_text(size=18, face="bold"), legend.text=element_text(size=16), legend.title=element_text(size=16))
```

# Extreme Responders ----------------------------------------------------------------------------------------------------------------------------------------

```{r}
# Now we look for extreme responders for each ion: defined as the individuals that show the highest/lowest normalized level of avg_nfoci at fluence 1.1 / dose 0.1, at 4h (not much repair yet)

Nextreme = 10 # Number of extreme low responders / high responders we want to consider

# For Fe:
fe_4h_dose1 = all_4h_norm[all_4h_norm$Rad_type == "Fe" & all_4h_norm$Fluence == "1.1",]
fe_4h_dose1_ranked = fe_4h_dose1[order(fe_4h_dose1$norm_nfoci, decreasing = TRUE),]
fe_4h_dose1_ranked2 = fe_4h_dose1[order(fe_4h_dose1$norm_nfoci),]

fe_high = fe_4h_dose1_ranked$Sample_ID[1:Nextreme] # Gives the list of Sample_ID for most sensitive
fe_low = fe_4h_dose1_ranked2$Sample_ID[1:Nextreme] # Gives the list of Sample_ID for most resistant

fe_4h_norm_extreme_high <- fe_avg_4h_norm[ which(fe_avg_4h_norm$Sample_ID %in% fe_high),] # Extract data from most sensitive only
fe_4h_norm_extreme_low <- fe_avg_4h_norm[ which(fe_avg_4h_norm$Sample_ID %in% fe_low),] # Extract data from most resistant only

colnames(fe_4h_norm_extreme_high)[colnames(fe_4h_norm_extreme_high) == "variable"] = "Fluence"
colnames(fe_4h_norm_extreme_high)[colnames(fe_4h_norm_extreme_high) == "value"] = "norm_nfoci"
colnames(fe_4h_norm_extreme_low)[colnames(fe_4h_norm_extreme_low) == "variable"] = "Fluence"
colnames(fe_4h_norm_extreme_low)[colnames(fe_4h_norm_extreme_low) == "value"] = "norm_nfoci"

fe_extreme_high = merge(fe_4h_norm_extreme_high, subject_meta, by.x = 'Sample_ID', by.y = 'Sample_ID', all.x = FALSE, all.y = FALSE)
col_trunc = c("Sample_ID","Fluence", "norm_nfoci","Age","Gender","BMI","BMI_group")
fe_extreme_high = fe_extreme_high[,colnames(fe_extreme_high) %in% col_trunc]
fe_extreme_high$BMI_group = factor(fe_extreme_high$BMI_group, levels = c("Normal", "Overweight", "Obese")) 

fe_extreme_low = merge(fe_4h_norm_extreme_low, subject_meta, by.x = 'Sample_ID', by.y = 'Sample_ID', all.x = FALSE, all.y = FALSE)
col_trunc = c("Sample_ID","Fluence", "norm_nfoci","Age","Gender","BMI","BMI_group")
fe_extreme_low = fe_extreme_low[,colnames(fe_extreme_low) %in% col_trunc]
fe_extreme_low$BMI_group = factor(fe_extreme_low$BMI_group, levels = c("Normal", "Overweight", "Obese")) 

# For Ar:
ar_4h_dose1 = all_4h_norm[all_4h_norm$Rad_type == "Ar" & all_4h_norm$Fluence == "1.1",]
ar_4h_dose1_ranked = ar_4h_dose1[order(ar_4h_dose1$norm_nfoci, decreasing = TRUE),]
ar_4h_dose1_ranked2 = ar_4h_dose1[order(ar_4h_dose1$norm_nfoci),]

ar_high = ar_4h_dose1_ranked$Sample_ID[1:Nextreme] # Gives the list of Sample_ID for most sensitive
ar_low = ar_4h_dose1_ranked2$Sample_ID[1:Nextreme] # Gives the list of Sample_ID for most resistant

ar_4h_norm_extreme_high <- ar_avg_4h_norm[ which(ar_avg_4h_norm$Sample_ID %in% ar_high),] # Extract data from most sensitive only
ar_4h_norm_extreme_low <- ar_avg_4h_norm[ which(ar_avg_4h_norm$Sample_ID %in% ar_low),] # Extract data from most resistant only

colnames(ar_4h_norm_extreme_high)[colnames(ar_4h_norm_extreme_high) == "variable"] = "Fluence"
colnames(ar_4h_norm_extreme_high)[colnames(ar_4h_norm_extreme_high) == "value"] = "norm_nfoci"
colnames(ar_4h_norm_extreme_low)[colnames(ar_4h_norm_extreme_low) == "variable"] = "Fluence"
colnames(ar_4h_norm_extreme_low)[colnames(ar_4h_norm_extreme_low) == "value"] = "norm_nfoci"

ar_extreme_high = merge(ar_4h_norm_extreme_high, subject_meta, by.x = 'Sample_ID', by.y = 'Sample_ID', all.x = FALSE, all.y = FALSE)
col_trunc = c("Sample_ID","Fluence", "norm_nfoci","Age","Gender","BMI","BMI_group")
ar_extreme_high = ar_extreme_high[,colnames(ar_extreme_high) %in% col_trunc]
ar_extreme_high$BMI_group = factor(ar_extreme_high$BMI_group, levels = c("Normal", "Overweight", "Obese")) 

ar_extreme_low = merge(ar_4h_norm_extreme_low, subject_meta, by.x = 'Sample_ID', by.y = 'Sample_ID', all.x = FALSE, all.y = FALSE)
col_trunc = c("Sample_ID","Fluence", "norm_nfoci","Age","Gender","BMI","BMI_group")
ar_extreme_low = ar_extreme_low[,colnames(ar_extreme_low) %in% col_trunc]
ar_extreme_low$BMI_group = factor(ar_extreme_low$BMI_group, levels = c("Normal", "Overweight", "Obese")) 


# For Si:
si_4h_dose1 = all_4h_norm[all_4h_norm$Rad_type == "Si" & all_4h_norm$Fluence == "1.1",]
si_4h_dose1_ranked = si_4h_dose1[order(si_4h_dose1$norm_nfoci, decreasing = TRUE),]
si_4h_dose1_ranked2 = si_4h_dose1[order(si_4h_dose1$norm_nfoci),]

si_high = si_4h_dose1_ranked$Sample_ID[1:Nextreme] # Gives the list of Sample_ID for most sensitive
si_low = si_4h_dose1_ranked2$Sample_ID[1:Nextreme] # Gives the list of Sample_ID for most resistant

si_4h_norm_extreme_high <- si_avg_4h_norm[ which(si_avg_4h_norm$Sample_ID %in% si_high),] # Extract data from most sensitive only
si_4h_norm_extreme_low <- si_avg_4h_norm[ which(si_avg_4h_norm$Sample_ID %in% si_low),] # Extract data from most resistant only

colnames(si_4h_norm_extreme_high)[colnames(si_4h_norm_extreme_high) == "variable"] = "Fluence"
colnames(si_4h_norm_extreme_high)[colnames(si_4h_norm_extreme_high) == "value"] = "norm_nfoci"
colnames(si_4h_norm_extreme_low)[colnames(si_4h_norm_extreme_low) == "variable"] = "Fluence"
colnames(si_4h_norm_extreme_low)[colnames(si_4h_norm_extreme_low) == "value"] = "norm_nfoci"

si_extreme_high = merge(si_4h_norm_extreme_high, subject_meta, by.x = 'Sample_ID', by.y = 'Sample_ID', all.x = FALSE, all.y = FALSE)
col_trunc = c("Sample_ID","Fluence", "norm_nfoci","Age","Gender","BMI","BMI_group")
si_extreme_high = si_extreme_high[,colnames(si_extreme_high) %in% col_trunc]
si_extreme_high$BMI_group = factor(si_extreme_high$BMI_group, levels = c("Normal", "Overweight", "Obese")) 

si_extreme_low = merge(si_4h_norm_extreme_low, subject_meta, by.x = 'Sample_ID', by.y = 'Sample_ID', all.x = FALSE, all.y = FALSE)
col_trunc = c("Sample_ID","Fluence", "norm_nfoci","Age","Gender","BMI","BMI_group")
si_extreme_low = si_extreme_low[,colnames(si_extreme_low) %in% col_trunc]
si_extreme_low$BMI_group = factor(si_extreme_low$BMI_group, levels = c("Normal", "Overweight", "Obese")) 


# For Gamma:
gamma_4h_dose1 = gamma_avg_4h_norm[gamma_avg_4h_norm$Dose == "0.1",]
gamma_4h_dose1_ranked = gamma_4h_dose1[order(gamma_4h_dose1$norm_nfoci, decreasing = TRUE),]
gamma_4h_dose1_ranked2 = gamma_4h_dose1[order(gamma_4h_dose1$norm_nfoci),]

gamma_high = gamma_4h_dose1_ranked$Sample_ID[1:Nextreme] # Gives the list of Sample_ID for most sensitive
gamma_low = gamma_4h_dose1_ranked2$Sample_ID[1:Nextreme] # Gives the list of Sample_ID for most resistant

gamma_4h_norm_extreme_high <- gamma_avg_4h_norm[ which(gamma_avg_4h_norm$Sample_ID %in% gamma_high),] # Extract data from most sensitive only
gamma_4h_norm_extreme_low <- gamma_avg_4h_norm[ which(gamma_avg_4h_norm$Sample_ID %in% gamma_low),] # Extract data from most resistant only

colnames(gamma_4h_norm_extreme_high)[colnames(gamma_4h_norm_extreme_high) == "variable"] = "Dose"
colnames(gamma_4h_norm_extreme_high)[colnames(gamma_4h_norm_extreme_high) == "value"] = "norm_nfoci"
colnames(gamma_4h_norm_extreme_low)[colnames(gamma_4h_norm_extreme_low) == "variable"] = "Dose"
colnames(gamma_4h_norm_extreme_low)[colnames(gamma_4h_norm_extreme_low) == "value"] = "norm_nfoci"

gamma_extreme_high = merge(gamma_4h_norm_extreme_high, subject_meta, by.x = 'Sample_ID', by.y = 'Sample_ID', all.x = FALSE, all.y = FALSE)
col_trunc = c("Sample_ID","Dose", "norm_nfoci","Age","Gender","BMI","BMI_group")
gamma_extreme_high = gamma_extreme_high[,colnames(gamma_extreme_high) %in% col_trunc]
gamma_extreme_high$BMI_group = factor(gamma_extreme_high$BMI_group, levels = c("Normal", "Overweight", "Obese")) 

gamma_extreme_low = merge(gamma_4h_norm_extreme_low, subject_meta, by.x = 'Sample_ID', by.y = 'Sample_ID', all.x = FALSE, all.y = FALSE)
col_trunc = c("Sample_ID","Dose", "norm_nfoci","Age","Gender","BMI","BMI_group")
gamma_extreme_low = gamma_extreme_low[,colnames(gamma_extreme_low) %in% col_trunc]
gamma_extreme_low$BMI_group = factor(gamma_extreme_low$BMI_group, levels = c("Normal", "Overweight", "Obese")) 
```

```{r}
# We plot the dose-response for these extreme responders

# For Fe
all_plot = ggplot(fe_extreme_high, aes(x = Fluence, y = norm_nfoci, col = Age, label = Sample_ID)) + geom_point(aes(shape=Gender), size=3) + geom_line(aes(group = Sample_ID, linetype = BMI_group), size=1) + ggtitle("Fe: top 10 sensitive")
all_plot + ylim(1,2.5) + xlab("Fluence (particles/100um^2)") + ylab("Normalized Average Number of Foci per Nucleus") + theme(axis.text.x = element_text(size=18)) + theme(axis.text.y = element_text(size=18)) + theme(axis.title.x = element_text(size=18)) + theme(axis.title.y = element_text(size=18)) + labs(col="Age") + theme(plot.title = element_text(size=18, face="bold"), legend.text=element_text(size=16), legend.title=element_text(size=16)) 

all_plot = ggplot(fe_extreme_low, aes(x = Fluence, y = norm_nfoci, col = Age, label = Sample_ID)) + geom_point(aes(shape=Gender), size=3) + geom_line(aes(group = Sample_ID, linetype = BMI_group), size=1) + ggtitle("Fe: top 10 resistant")
all_plot + xlab("Fluence (particles/100um^2)") + ylab("Normalized Average Number of Foci per Nucleus") + theme(axis.text.x = element_text(size=18)) + theme(axis.text.y = element_text(size=18)) + theme(axis.title.x = element_text(size=18)) + theme(axis.title.y = element_text(size=18)) + labs(col="Age") + theme(plot.title = element_text(size=18, face="bold"), legend.text=element_text(size=16), legend.title=element_text(size=16)) 

# For Ar
all_plot = ggplot(ar_extreme_high, aes(x = Fluence, y = norm_nfoci, col = Age, label = Sample_ID)) + geom_point(aes(shape=Gender), size=3) + geom_line(aes(group = Sample_ID, linetype = BMI_group), size=1) + ggtitle("Ar: top 10 sensitive")
all_plot + ylim(1,2.5) + xlab("Fluence (particles/100um^2)") + ylab("Normalized Average Number of Foci per Nucleus") + theme(axis.text.x = element_text(size=18)) + theme(axis.text.y = element_text(size=18)) + theme(axis.title.x = element_text(size=18)) + theme(axis.title.y = element_text(size=18)) + labs(col="Age") + theme(plot.title = element_text(size=18, face="bold"), legend.text=element_text(size=16), legend.title=element_text(size=16)) 

all_plot = ggplot(ar_extreme_low, aes(x = Fluence, y = norm_nfoci, col = Age, label = Sample_ID)) + geom_point(aes(shape=Gender), size=3) + geom_line(aes(group = Sample_ID, linetype = BMI_group), size=1) + ggtitle("Ar: top 10 resistant")
all_plot + xlab("Fluence (particles/100um^2)") + ylab("Normalized Average Number of Foci per Nucleus") + theme(axis.text.x = element_text(size=18)) + theme(axis.text.y = element_text(size=18)) + theme(axis.title.x = element_text(size=18)) + theme(axis.title.y = element_text(size=18)) + labs(col="Age") + theme(plot.title = element_text(size=18, face="bold"), legend.text=element_text(size=16), legend.title=element_text(size=16)) 

# For Si
all_plot = ggplot(si_extreme_high, aes(x = Fluence, y = norm_nfoci, col = Age, label = Sample_ID)) + geom_point(aes(shape=Gender), size=3) + geom_line(aes(group = Sample_ID, linetype = BMI_group), size=1) + ggtitle("Si: top 10 sensitive")
all_plot + ylim(1,2.5) + xlab("Fluence (particles/100um^2)") + ylab("Normalized Average Number of Foci per Nucleus") + theme(axis.text.x = element_text(size=18)) + theme(axis.text.y = element_text(size=18)) + theme(axis.title.x = element_text(size=18)) + theme(axis.title.y = element_text(size=18)) + labs(col="Age") + theme(plot.title = element_text(size=18, face="bold"), legend.text=element_text(size=16), legend.title=element_text(size=16)) 

all_plot = ggplot(si_extreme_low, aes(x = Fluence, y = norm_nfoci, col = Age, label = Sample_ID)) + geom_point(aes(shape=Gender), size=3) + geom_line(aes(group = Sample_ID, linetype = BMI_group), size=1) + ggtitle("Si: top 10 resistant")
all_plot + xlab("Fluence (particles/100um^2)") + ylab("Normalized Average Number of Foci per Nucleus") + theme(axis.text.x = element_text(size=18)) + theme(axis.text.y = element_text(size=18)) + theme(axis.title.x = element_text(size=18)) + theme(axis.title.y = element_text(size=18)) + labs(col="Age") + theme(plot.title = element_text(size=18, face="bold"), legend.text=element_text(size=16), legend.title=element_text(size=16)) 

# For Gamma
all_plot = ggplot(gamma_extreme_high, aes(x = Dose, y = norm_nfoci, col = Age, label = Sample_ID)) + geom_point(aes(shape=Gender), size=3) + geom_line(aes(group = Sample_ID, linetype = BMI_group), size=1) + ggtitle("Gamma: top 10 sensitive")
all_plot + xlab("Dose (Gy)") + ylab("Normalized Average Number of Foci per Nucleus") + theme(axis.text.x = element_text(size=18)) + theme(axis.text.y = element_text(size=18)) + theme(axis.title.x = element_text(size=18)) + theme(axis.title.y = element_text(size=18)) + labs(col="Age") + theme(plot.title = element_text(size=18, face="bold"), legend.text=element_text(size=16), legend.title=element_text(size=16)) 

all_plot = ggplot(gamma_extreme_low, aes(x = Dose, y = norm_nfoci, col = Age, label = Sample_ID)) + geom_point(aes(shape=Gender), size=3) + geom_line(aes(group = Sample_ID, linetype = BMI_group), size=1) + ggtitle("Gamma: top 10 resistant")
all_plot + xlab("Dose (Gy)") + ylab("Normalized Average Number of Foci per Nucleus") + theme(axis.text.x = element_text(size=18)) + theme(axis.text.y = element_text(size=18)) + theme(axis.title.x = element_text(size=18)) + theme(axis.title.y = element_text(size=18)) + labs(col="Age") + theme(plot.title = element_text(size=18, face="bold"), legend.text=element_text(size=16), legend.title=element_text(size=16))
```

```{r}
# Another way to define extreme responders: Determine extreme responders based on bottom and top quantiles for normalized number of foci at first fluence/dose

quant_fe_bottom = quantile(fe_avg_4h_norm$value, na.rm = TRUE)[2]
quant_fe_top = quantile(fe_avg_4h_norm$value, na.rm = TRUE)[4]
subj_fe_res_bottom = fe_avg_4h_norm$Sample_ID[fe_avg_4h_norm$value < quant_fe_bottom]
subj_fe_res_top = fe_avg_4h_norm$Sample_ID[fe_avg_4h_norm$value > quant_fe_bottom]

fe_res_bottom <- fe_avg_4h_norm[ which(fe_avg_4h_norm$Sample_ID %in% subj_fe_res_bottom),]
fe_res_bottom = merge(fe_res_bottom, subject_meta, by.x = 'Sample_ID', by.y = 'Sample_ID', all.x = FALSE, all.y = FALSE) # we want to add sample metadata
col_trunc = c("Sample_ID","Fluence", "norm_nfoci","Age","Gender","BMI","BMI_group")
fe_res_bottom = fe_res_bottom[,colnames(fe_res_bottom) %in% col_trunc]
fe_res_bottom$BMI_group = factor(fe_res_bottom$BMI_group, levels = c("Normal", "Overweight", "Obese"))
hist(fe_res_bottom$Age)

quant_si_bottom = quantile(si_avg_4h_norm$value, na.rm = TRUE)[2]
quant_si_top = quantile(si_avg_4h_norm$value, na.rm = TRUE)[4]
subj_si_res_bottom = si_avg_4h_norm$Sample_ID[si_avg_4h_norm$value < quant_si_bottom]
subj_si_res_top = si_avg_4h_norm$Sample_ID[si_avg_4h_norm$value > quant_si_bottom]

quant_ar_bottom = quantile(ar_avg_4h_norm$value, na.rm = TRUE)[2]
quant_ar_top = quantile(ar_avg_4h_norm$value, na.rm = TRUE)[4]
subj_ar_res_bottom = ar_avg_4h_norm$Sample_ID[ar_avg_4h_norm$value < quant_ar_bottom]
subj_ar_res_top = ar_avg_4h_norm$Sample_ID[ar_avg_4h_norm$value > quant_ar_bottom]

quant_gamma_bottom = quantile(gamma_avg_4h_norm$value, na.rm = TRUE)[2]
quant_gamma_top = quantile(gamma_avg_4h_norm$value, na.rm = TRUE)[4]
subj_gamma_res_bottom = gamma_avg_4h_norm$Sample_ID[gamma_avg_4h_norm$value < quant_gamma_bottom]
subj_gamma_res_top = gamma_avg_4h_norm$Sample_ID[gamma_avg_4h_norm$value > quant_gamma_bottom]

subj_common_res_bottom = intersect(intersect(subj_fe_res_bottom, subj_si_res_bottom), subj_ar_res_bottom)
subj_common_res_top = intersect(intersect(subj_fe_res_top, subj_si_res_top),subj_ar_res_top)
```

# Extreme Baselines -------------------------------------------------------------------------------------------------------------------------------------------

```{r}
N_extreme = 2*10 # Number of individuals we want to consider for each group of extreme baselines

# For Fe:
fe_avg_4h_norm_base = merge(fe_avg_4h_norm,baseline_avgfoci, by.x = 'Sample_ID', by.y = 'Sample_ID', all.x = FALSE, all.y = FALSE) #adds baseline data
colnames(fe_avg_4h_norm_base)[colnames(fe_avg_4h_norm_base) == "variable"] = "Fluence"
colnames(fe_avg_4h_norm_base)[colnames(fe_avg_4h_norm_base) == "avg_nfoci"] = "baseline_avg_nfoci"
colnames(fe_avg_4h_norm_base)[colnames(fe_avg_4h_norm_base) == "value"] = "avg_nfoci"

fe_baseline_order_min <- fe_avg_4h_norm_base[order(fe_avg_4h_norm_base$baseline_avg_nfoci),]
fe_baseline_order_max <- fe_avg_4h_norm_base[order(fe_avg_4h_norm_base$baseline_avg_nfoci, decreasing = TRUE),]
fe_baseline_low <- fe_baseline_order_min[1:N_extreme,]
fe_baseline_low["Baseline"] <- rep("low",nrow(fe_baseline_low))
fe_baseline_high <- fe_baseline_order_max[1:N_extreme,]
fe_baseline_high["Baseline"] <- rep("high",nrow(fe_baseline_high))
fe_baseline_extreme <- rbind.data.frame(fe_baseline_low,fe_baseline_high)

plot_fe_extreme_base = ggplot(fe_baseline_extreme, aes(x = Fluence, y = avg_nfoci, fill=Baseline)) +  
  geom_boxplot(lwd=1) + scale_fill_brewer(palette = "Pastel2") + ggtitle("Iron")
plot_fe_extreme_base + ylim(1,2.2) + xlab("Fluence (particles/100um^2)") + ylab("Normalized Average Number of Foci per Nucleus") + theme(axis.text.x = element_text(size=18)) + theme(axis.text.y = element_text(size=18)) + theme(axis.title.x = element_text(size=18)) + theme(axis.title.y = element_text(size=18)) + labs(col="Radiation") + theme(plot.title = element_text(size=18, face="bold"), legend.text=element_text(size=16), legend.title=element_text(size=16))

# For Ar:
ar_avg_4h_norm_base = merge(ar_avg_4h_norm,baseline_avgfoci, by.x = 'Sample_ID', by.y = 'Sample_ID', all.x = FALSE, all.y = FALSE)
colnames(ar_avg_4h_norm_base)[colnames(ar_avg_4h_norm_base) == "variable"] = "Fluence"
colnames(ar_avg_4h_norm_base)[colnames(ar_avg_4h_norm_base) == "avg_nfoci"] = "baseline_avg_nfoci"
colnames(ar_avg_4h_norm_base)[colnames(ar_avg_4h_norm_base) == "value"] = "avg_nfoci"

ar_baseline_order_min <- ar_avg_4h_norm_base[order(ar_avg_4h_norm_base$baseline_avg_nfoci),]
ar_baseline_order_max <- ar_avg_4h_norm_base[order(ar_avg_4h_norm_base$baseline_avg_nfoci, decreasing = TRUE),]
ar_baseline_low <- ar_baseline_order_min[1:N_extreme,]
ar_baseline_low["Baseline"] <- rep("low",nrow(ar_baseline_low))
ar_baseline_high <- ar_baseline_order_max[1:N_extreme,]
ar_baseline_high["Baseline"] <- rep("high",nrow(ar_baseline_high))
ar_baseline_extreme <- rbind.data.frame(ar_baseline_low,ar_baseline_high)

plot_ar_extreme_base = ggplot(ar_baseline_extreme, aes(x = Fluence, y = avg_nfoci, fill=Baseline)) +  
  geom_boxplot(lwd=1) + scale_fill_brewer(palette = "Pastel2") + ggtitle("Argon")
plot_ar_extreme_base + ylim(1,2.2) + xlab("Fluence (particles/100um^2)") + ylab("Normalized Average Number of Foci per Nucleus") + theme(axis.text.x = element_text(size=18)) + theme(axis.text.y = element_text(size=18)) + theme(axis.title.x = element_text(size=18)) + theme(axis.title.y = element_text(size=18)) + labs(col="Radiation") + theme(plot.title = element_text(size=18, face="bold"), legend.text=element_text(size=16), legend.title=element_text(size=16))

# For Si:
si_avg_4h_norm_base = merge(si_avg_4h_norm,baseline_avgfoci, by.x = 'Sample_ID', by.y = 'Sample_ID', all.x = FALSE, all.y = FALSE)
colnames(si_avg_4h_norm_base)[colnames(si_avg_4h_norm_base) == "variable"] = "Fluence"
colnames(si_avg_4h_norm_base)[colnames(si_avg_4h_norm_base) == "avg_nfoci"] = "baseline_avg_nfoci"
colnames(si_avg_4h_norm_base)[colnames(si_avg_4h_norm_base) == "value"] = "avg_nfoci"

si_baseline_order_min <- si_avg_4h_norm_base[order(si_avg_4h_norm_base$baseline_avg_nfoci),]
si_baseline_order_max <- si_avg_4h_norm_base[order(si_avg_4h_norm_base$baseline_avg_nfoci, decreasing = TRUE),]
si_baseline_low <- si_baseline_order_min[1:N_extreme,]
si_baseline_low["Baseline"] <- rep("low",nrow(si_baseline_low))
si_baseline_high <- si_baseline_order_max[1:N_extreme,]
si_baseline_high["Baseline"] <- rep("high",nrow(si_baseline_high))
si_baseline_extreme <- rbind.data.frame(si_baseline_low,si_baseline_high)

plot_si_extreme_base = ggplot(si_baseline_extreme, aes(x = Fluence, y = avg_nfoci, fill=Baseline)) +  
  geom_boxplot(lwd=1) + scale_fill_brewer(palette = "Pastel2") + ggtitle("Silicon")
plot_si_extreme_base + ylim(1,2) + xlab("Fluence (particles/100um^2)") + ylab("Normalized Average Number of Foci per Nucleus") + theme(axis.text.x = element_text(size=18)) + theme(axis.text.y = element_text(size=18)) + theme(axis.title.x = element_text(size=18)) + theme(axis.title.y = element_text(size=18)) + labs(col="Radiation") + theme(plot.title = element_text(size=18, face="bold"), legend.text=element_text(size=16), legend.title=element_text(size=16))

# For Gamma:
gamma_avg_4h_norm_base = merge(gamma_avg_4h_norm,baseline_avgfoci, by.x = 'Sample_ID', by.y = 'Sample_ID', all.x = FALSE, all.y = FALSE)
colnames(gamma_avg_4h_norm_base)[colnames(gamma_avg_4h_norm_base) == "avg_nfoci"] = "baseline_avg_nfoci"

gamma_baseline_order_min <- gamma_avg_4h_norm_base[order(gamma_avg_4h_norm_base$baseline_avg_nfoci),]
gamma_baseline_order_max <- gamma_avg_4h_norm_base[order(gamma_avg_4h_norm_base$baseline_avg_nfoci, decreasing = TRUE),]
gamma_baseline_low <- gamma_baseline_order_min[1:N_extreme,]
gamma_baseline_low["Baseline"] <- rep("low",nrow(gamma_baseline_low))
gamma_baseline_high <- gamma_baseline_order_max[1:N_extreme,]
gamma_baseline_high["Baseline"] <- rep("high",nrow(gamma_baseline_high))
gamma_baseline_extreme <- rbind.data.frame(gamma_baseline_low,gamma_baseline_high)

plot_gamma_extreme_base = ggplot(gamma_baseline_extreme, aes(x = Dose, y = norm_nfoci, fill=Baseline)) +  
  geom_boxplot(lwd=1) + scale_fill_brewer(palette = "Pastel2") + ggtitle("Gamma")
plot_gamma_extreme_base + ylim(1,2) + xlab("Dose (Gy)") + ylab("Normalized Average Number of Foci per Nucleus") + theme(axis.text.x = element_text(size=18)) + theme(axis.text.y = element_text(size=18)) + theme(axis.title.x = element_text(size=18)) + theme(axis.title.y = element_text(size=18)) + labs(col="Radiation") + theme(plot.title = element_text(size=18, face="bold"), legend.text=element_text(size=16), legend.title=element_text(size=16))
```